---
title: "Playing with AMOVA results"
author: "Eric D. Crandall"
date: "November 4, 2015"
output: html_document
---
# Load the AMOVA results
And the trimmed database (IP data only)
```{r}
load("~/github/popgenDB/output/FST_table_amova_output.R")
load("~/github/popgenDB/output/ipdb_ip.R")
hypotheses<-c("ECOREGION", "PROVINCE","REALM","Bowen","Keith","Kulbicki_r","Kulbicki_b", "VeronDivis")

```

# Plot Mean FCTs

```{r}
# melt the list and pull out the FCT values
library(reshape2)
melted<-melt(amovastats)
FCT<-melted[which(melted$variable=="FCT"),]
FSC<-melted[which(melted$variable=="FSC"),]
FST<-melted[which(melted$variable=="FST"),]

# Calculate the number of regions in each hypothesis
hyp_ip<-ipdb_ip[,names(ipdb_ip) %in% hypotheses]
k<-sapply(X=hyp_ip,FUN= function(x) length(unique(x))) 
k<-k[order(match(names(k),hypotheses))] # reorder this vector to match the order in the hypotheses vector
k

## get the means
by(data=FCT$value,INDICES =as.factor(FCT$L1),FUN=mean, na.rm=T)
```

## Plot them in ggplot2

```{r}
library(ggplot2)
plot<-ggplot(data=FCT, aes(x=factor(L1,hypotheses),y=value))
plot<-plot + geom_boxplot(notch=T) + ylim(c(-0.1,1))+
  labs(x="Hypothesis", y=expression('F'[CT]), title="Regional Indo-Pacific Structure explained by Biogeographic Hypotheses") +
  scale_x_discrete(label=paste(names(k),"\nk = ",k))
plot
```

# Make pivot tables of each regionalization

```{r, eval=F}
ipdb_e<-dcast(ipdb_ip, Genus_species_locus ~ VeronDivis, fun.aggregate=length)
write.xlsx(x=ipdb_e, file="Regionalizations.xlsx",sheetName="VeronDivis",append=T)
```

# Measure support for each hypothesis

Get the values for each hypothesis for a given criterion - here I use BIC - and rank them for each species, then choose the "best" hypothesis for each species based on the criterion.
```{r}
criterion<-"BIC"
# find the maximum number of species from any of the 8 hypotheses
maxlength<-max(sapply(amovastats,function(x) length(x[[1]]))) 
# create an empty data frame with row names from the hypothesis with the most values
crit_df<-data.frame(setNames(replicate(maxlength,numeric(0), simplify = F),nm=row.names(amovastats[["PROVINCE"]]))) 

#Loop through the hypotheses, pulling out the values for criterion,transpose it, and then merge these values into the dataframe from the previous hypothesis 
for(h in names(amovastats)){
  crit_df<-merge(crit_df,t(amovastats[[h]][criterion]),all=T,sort=F)
}
#get the hypothesis names in there
row.names(crit_df)<-names(amovastats)

#rank the hypotheses for each species
crit_rank<-as.data.frame(sapply(crit_df,rank,na.last="keep",ties.method="average"))
row.names(crit_rank)<-names(amovastats)

## remove gsls with more than 3 missing models
crit_rank<-crit_rank[, colSums(is.na(crit_rank)) < nrow(crit_rank)-5]  

## choose the best hypothesis or set of hypotheses for each species  
best_hypothesis<-sapply(crit_rank,function(x){row.names(crit_rank)[which(x==min(x,na.rm=T)) ]})
```

## Make a bar graph of best support for each hypothesis among species
```{r}
barplot<-ggplot(data=as.data.frame(unlist(best_hypothesis)), aes(x=unlist(best_hypothesis),y=(..count..)/sum(..count..) ))
barplot<-barplot+geom_histogram() + labs(x="Hypothesis",y="Proportion of Species", title="Proportional Support for Biogeographic Hypotheses based on BIC")
barplot
```

## Calculate relative probability from Johnson and Omland 2004
```{r}
# lookup the minimum BIC value for each species (which.min works better here, because it returns only the first instance of the minimum value)
minBIC<-sapply(crit_df,function(x){x[which.min(x)]}) 

#J&O box 4 eqn 1. scale() seems to be the way to go here, using the minBIC as the centering vector
crit_df_deltaI<-scale(crit_df, center=minBIC, scale=F) 

#J&O box 4 eqn 4. numerator, plus make it a data frame
crit_df_deltaI_b<-as.data.frame(exp(-0.5*crit_df_deltaI)) 

#J&O box 4 eqn 4. denominator
crit_df_deltaI_sums<-sapply(crit_df_deltaI_b,sum,na.rm=T) 

# this time use the scale argument of scale to divide each column by the corresponding sum
crit_df_relative_prob<-scale(crit_df_deltaI_b,center=F,scale=crit_df_deltaI_sums) 
```

## Make a heatmap of relative probability of each hypothesis
```{r}
#melt for ggplot2
relprob<-melt(crit_df_relative_prob)
colnames(relprob)<-c("Hypothesis","Species","Relative_Probability")

#baseplot
rp<-ggplot(relprob,aes(y=Species,x=Hypothesis,fill=Relative_Probability))

#add geom_tile, turn the x-axis elements by 90 degrees, reverse the names on the y-axis, and use a diverging color scheme to highlight hypotheses with >50% rel prob.
rp<-rp+geom_tile()+
  theme(axis.text.x = element_text(angle = 90, hjust = 1))+
  ylim(rev(levels(relprob$Species)))+
  scale_fill_gradient2(low = "blue", mid = "white",
                       high = "red", midpoint = 0.5, space = "rgb",
                       na.value = "grey50", guide = "colourbar")
rp
```
